---
title: "Using {utVizSunburst}"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using {utVizSunburst}}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(utVizSunburst)
```

## An example dataset

In the utVizSunburst package, and example dataset "admissions" is available.

```{r}
data(admissions, package = "utVizSunburst")

head(admissions)
```

The dataset contains `r ncol(admissions)` variables and `r nrow(admissions)` observations.

## A simple sunburst chart

The function `sunburst()` generates a sunburst chart. The two main arguments to this function are:

- `data` (which provides the data-frame for the chart);
- and `steps` (indicating which columns of that data-frame should be presented in the chart).

```{r}
sunburst(
  data = admissions,
  steps = c("student_type", "outcome")
)
```

The sectors that are highlighted when the user hovers over the chart are proportional to the
percentage of individuals that make up that sector.

Note that the order of the `steps` values _does_ matter. It defines the order that each variable is
presented by the chart from inner-most ring (the first entry) to outer-most ring (the final entry).

In the above, `student_type` was on the inside, because it was the first entry in `steps`. But if we
reverse the order, then `outcome` becomes the inner-most ring.

```{r}
sunburst(
  data = admissions,
  steps = c("outcome", "student_type")
)
```

When creating a sunburst chart, any number of the columns in `data` can be used in the `steps`
argument. As such, we can make a chart containing just a single ring:

```{r}
sunburst(admissions, steps = "ipeds_race_ethn")
```

Or we can make a chart containing many rings:

```{r}
sunburst(
  admissions,
  steps = c("gpa", "club", "student_type", "outcome", "college")
)
```

# Sizing the chart

The initial size of a sunburst chart can be specified using the `width` and `height` arguments in
`sunburst()`. These take any CSS sizing format.

```{r}
sunburst(admissions, c("student_type", "outcome"), width = "200px")
```

```{r}
sunburst(admissions, c("student_type", "outcome"), width = "100%", height = "100%")
```

## User interaction

Each row of `data` corresponds to an observation. In a university admissions setting, that might be
information for a single student. The sectors of the sunburst chart represent a proportion of
these observations. The user can interact with a chart by hovering or clicking on sectors, giving a
way to gain further information about the data.

### Hovering

Hovering over a sector of the sunburst chart highlights that sector and sectors lying on the path
leading up to that sector. Paths in the chart go outwards from the inner-most ring and, while
hovering over a sector, the level of each variable along that path is indicated in text above the
chart (the breadcrumb trail for the path). Also, the proportion of individuals that are within a
given sector is displayed in the center of the chart.

### Clicking

Clicking on a sector expands that sector to fill the whole chart. Zooming, or drilling-down, into
the chart in this way allows for a deeper analysis of specific subsets of the original data. On
zooming, the subset of the original chart that lies outside of the clicked sector is presented. This
only works for sectors that are not in the outer ring of the chart (since there is no way to
partition a sector in the outer ring of the chart further).

There is a reset button, which allows a user to revert back to the initial chart after they have
drilled-down into specific subsets of the data.

## Colors

By default, the sector-colors in the sunburst chart are based on the color of the inner-most ring.
A palette of colors can be supplied by the user if the default colors need to change. These are used
to color all the sectors of the inner ring in order from largest to smallest.

```{r}
unique(admissions$outcome)

# The colors from RColorBrewer::brewer.pal(4, "RdBu")
palette = c("#CA0020", "#F4A582", "#92C5DE", "#0571B0")

sunburst(
  admissions,
  steps = c("outcome", "student_type"),
  palette = palette
)
```

The named colors provided by R can be used in the `palette` argument:

```{r}
sunburst(
  admissions,
  steps = c("outcome", "student_type"),
  palette = c("firebrick", "goldenrod", "dodgerblue", "dimgrey")
)
```
